---
title: "Replicating Mondak & Sanders (2003): Tolerance and Intolerance, 1976–1998"
subtitle: "DWV Assignment 6"
author: "Josh Arden"
date: today
format:
  pdf:
    toc: false
    number-sections: false
    fig-width: 7
    fig-height: 4.5
    fig-pos: "H"
  html: default
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# Runs all data cleaning and builds every figure object from the script.
# ps6.R and gss7224_r1.dta must be in the same folder as this .qmd file.
source("ps6.R")

library(knitr)

# Shared labels for the replication figures
rep_x <- "Survey year"
```

All code used to clean the data and build the figures is in `ps6.R` in the
accompanying GitHub repository: https://github.com/josharden-jpa/DWV_PS6.

# Replication

## Figure 1: The 0–15 Tolerance Battery

```{r}
#| label: fig-1
#| fig-cap: "Mean score on the 15-item GSS modified Stouffer battery, 1976–1998. Higher scores represent greater intolerance."

fig1 +
  labs(title    = "The 0–15 GSS Modified Stouffer Tolerance Battery, 1976–1998",
       subtitle = "High scores represent intolerance",
       x = rep_x, y = "Mean intolerance score (0–15)",
       caption  = gss_caption) +
  theme_mondak
```

## Figure 2: Intolerance Toward Racists and Homosexuals

```{r}
#| label: fig-2
#| fig-cap: "Mean score on the three tolerance items for each target group, 1976–1998 (0 = fully tolerant, 3 = fully intolerant)."

fig2 +
  labs(title    = "Intolerance Toward Racists and Homosexuals, 1976–1998",
       subtitle = "Blue = racists; red = homosexuals",
       x = rep_x, y = "Mean intolerance score (0–3)",
       caption  = gss_caption) +
  theme_mondak
```

## Figure 5: A Dichotomous Measure of Intolerance

```{r}
#| label: fig-5
#| fig-cap: "Share of respondents giving at least one intolerant answer on the 15-item battery, 1976–1998."

fig5 +
  labs(title    = "A Dichotomous Measure of Tolerance and Intolerance, 1976–1998",
       subtitle = "Intolerant = any score above 0 out of 15",
       x = rep_x, y = "Proportion intolerant",
       caption  = gss_caption) +
  theme_mondak
```

## Extension: All Available Years

```{r}
#| label: fig-1-all
#| fig-cap: "Figure 1 extended through 2024."

fig1_all +
  labs(title = "The 0–15 Tolerance Battery, All Years",
       subtitle = "High scores represent intolerance",
       x = rep_x, y = "Mean intolerance score (0–15)",
       caption = gss_caption) +
  theme_mondak
```

```{r}
#| label: fig-2-all
#| fig-cap: "Figure 2 extended through 2024."

fig2_all +
  labs(title = "Intolerance Toward Racists and Homosexuals, All Years",
       subtitle = "Blue = racists; red = homosexuals",
       x = rep_x, y = "Mean intolerance score (0–3)",
       caption = gss_caption) +
  theme_mondak
```

```{r}
#| label: fig-5-all
#| fig-cap: "Figure 5 extended through 2024."

fig5_all +
  labs(title = "Dichotomous Measure of Intolerance, All Years",
       subtitle = "Intolerant = any score above 0 out of 15",
       x = rep_x, y = "Proportion intolerant",
       caption = gss_caption) +
  theme_mondak
```

# Original Work: Two GSS Variables in 2024

## Television Viewing (TVHOURS)

```{r}
#| label: tbl-tv
#| tbl-cap: "Summary statistics for daily hours of television, 2024 GSS."

kable(tibble(
  N      = nrow(tv),
  Mean   = round(mean(tv$tvhours), 2),
  Median = median(tv$tvhours),
  SD     = round(sd(tv$tvhours), 2),
  `25th pct.` = quantile(tv$tvhours, 0.25),
  `75th pct.` = quantile(tv$tvhours, 0.75),
  Max    = max(tv$tvhours)
))
```

```{r}
#| label: fig-tv
#| fig-cap: "Distribution of self-reported daily television viewing, 2024 GSS."

fig_tv
```

The most common answer is two hours, given by about 500 respondents, and the
median is also two hours. The mean, however, is a little over three hours.
The middle 50% of respondents watch between one and four hours a day. The
mean is pulled upward by outliers who report watching 20 or 24 hours a day,
which probably means they literally never turn their TV off, even while they
are sleeping. The distribution is skewed right: most answers bunch up around
one, two, and three hours, with a small tail stretching out to the right.

## Going to a Bar or Tavern (SOCBAR)

```{r}
#| label: fig-bar
#| fig-cap: "How often respondents go to a bar or tavern, 2024 GSS."

fig_bar
```

The median answer is "about once a year," which is quite surprising to me,
and 45% of respondents never go to a bar or tavern at all. I can understand
never going, but I would like to know what the occasion is for the people who
go once a year. Maybe it is for their own birthday, or for New Year's, the one
night out with friends. Apart from a small bump from "about once a year" to
"several times a year," the number of respondents decreases as the frequency
increases. In other words, the more trips to the bar an answer represents, the
fewer respondents give that answer. The middle 50% of respondents fall
between "about once a month" and "never." Technically, that range covers four
of the seven answer categories, but in substance the answers are not far
apart: everyone in the middle half goes to a bar once a month at most.

# Discussion

## 1. Problems With the Replication

The main problems I encountered in this replication were figuring out how the
responses were coded and how to properly recode them, and working out whether
each score should run from 0 to 15 or from 0 to 3. It was also tricky to
determine exactly what kind of graph was being used in the original figures.

There was also a bug in my first version of Figure 2. When calculating the
scores for intolerance toward homosexuals, I used the wrong data object (the
one that had not been filtered to 1976–1998 or recoded). Because none of the
recoded columns existed in that object, every respondent got a score of 0, and
the line for homosexuals showed up flat at zero across all years. Once I used
the correctly filtered and recoded data, the line matched the paper's
downward trend.

## 2. Interpreting Figures 1, 2, and 5

In Figure 1, the mean intolerance score drops by roughly 27% between 1977
and 1998. This doesn't tell us about any one particular group: some answers
could be becoming more tolerant while others become less tolerant. Overall,
though, people are scoring as more tolerant than they were decades ago.

Figure 2 shows that while intolerance toward racists stays relatively the
same, intolerance toward homosexuals declines. So even though people don't
feel any better or worse about racists, they do become more tolerant toward
homosexuals over the same period. In other words, it isn't that people
overall are disliking bigotry more; by this scoring, they are presumably
becoming more accepting of homosexuality. This changes how Figure 1 should be
read: it could be that only tolerance toward the specific groups named in the
questions has risen, while overall intolerance, especially toward new ideas
and groups beyond the ones the questions were about at the time, has not.

Figure 5 shows that the proportion of people who were intolerant in any way
has gone down. About 15% of respondents answered tolerant to all of the
questions in 1988, compared with 21% in 1998. Even so, a significant portion
of respondents, roughly four in five, remain intolerant of at least one group.

## 3. What Has Changed Since 1998?

Many things have changed since 1998. Looking at intolerance toward
homosexuals, for example, same-sex marriage is now legal nationwide. This
could show that attitudes were already trending toward acceptance, or the
newer norms, standards, and laws might in turn lead people to become more
tolerant because same-sex relationships are more normalized. There has also
been a lot of division since 1998, and many events could have a big effect on
how people answer the tolerance questions. These include 9/11, which caused a
lot of intolerance toward Muslim and Arab people. So some of the scores might
be different for different reasons.
